Impact offsetting

### 2022-02-07 paper.tex
First draft of stochastic baselines

	\appendix
	\section{Non-Markovian rewards}
	 In the main text I pretended everything is Markovian, but at least Krakovna's original approach
		is not. A definition that would actually apply to her appraoch is presented below,
		but was too unwieldy to work with and did not actually provide much insight.
		\begin{definition}[Side effect avoidance]
			\emph{Side effect avoidance} is a class of approaches for mitigating reward misspecification, which
			modify the reward passed to the agent according to the equation
			\begin{equation}
				\modifiedRewardFunction\left( \left( s_i \right)_{i=0}^t \right) =
				\corruption\left( s_t \right) -
				\lambda \cdot \deviationFromBaseline\left( \left( s_i \right)_{i=0}^t, \baseline\left( \left( s_i \right)_{i=0}^t \right) \right),
				\label{eq:sideEffectAvoidance-nonMarkov}
			\end{equation}
			where $t$ is the time step, $s_i$ is the state at time $i$, $\baseline$ is a probability
			distribution of trajectories of the baseline agent, and $\deviationFromBaseline$ is a measure
			of the deviation between a trajectory and a baseline distribution of trajectories.

>>>>

\section{Stochastic side-effect avoidance}
	This section presents an as of yet unexplored choice for $\baseline$.
	\subsection{Basic idea}
	 We would like to answer the counterfactual question ``What would the world look like if we used
		the baseline policy rather than the policy we actually used?''.
	 A solution to this problem would be a generator of baseline states that would,
		given an arbitrary state $s$, produce a distribution of states that could have
		been encountered instead of $s$.

		There is an obvious degree of freedom in the counterfactual question,
		even before we get to specifying what the generator should be doing.
		When should the counterfactual intervention happen? The answer to this
		question is the main difference between Krakovna's and Turner's approaches
		-- Krakovna answers ``at the beginning of time'', while Turner ``just a step ago''.
		Both these answers introduce some problems. I would argue we should be looking at
		a discounted combination of counterfactual interventions, e.g. rollout using the baseline
		policy starting at $4^n$ steps ago for $n \in \left\{ 0, \dots, k \right\}$.
		We should discount the penalties related to these counterfactuals by $2^k$.
		This avoids the tractability issues with Krakovna's approach, while avoiding the problem
		with Turner immediately accepting any loss in reversability after the agent is one step away
		form the decision that caused it.

		(I have a brilliant thought about this corresponding to the sunk cost fallacy, but
		not enough energy to write this out properly. Soemthingsomething
		only caring about the last decision makes you miss that your whole policy got you
		into deep trouble.)

		The generator itself could be trained on random policies with baseline rollouts
		computed starting from past states. Note that this requires the baseline to
		necessarily be a complete policy, being able to act in any state. This might
		be the inaction policy, but that would likely be unsatisfactory as a safe choice
		in general.
	\subsection{CVAE}
		A relatively natural approach to the above problem is to use a slightly modified CVAE.
		We create a rollout of a random policy $\pi$, of the form $\left( s_0, \dots s_T \right)$.
		We also create rollouts of the baseline policy $\pi_b$ of the forms
		\begin{equation*}
			\left( s^{\left( m \right)}_0 = s_m, \dots, s^{\left( m \right)}_{T-m} \right).
		\end{equation*}
		For every $t$ we use $s_t$ as an input to the CVAE, backpropagating a combination of
		$k$ losses
		\begin{equation*}
			L(s_t, s^{\left( t-1 \right)}_{1}, s^{\left( t-4 \right)}_4, \dots, s^{\left( t-4^k \right)}_{4^k}) = \sum_{l = 0}^k 2^{-l}L_{CVAE}\left( s_t, s^{\left( t-4^l \right)}_{4^l} \right)
		\end{equation*}
		Where $L_{CVAE}$ is the standard CVAE loss.

		(It might be preferable to actually train $k$ different models, but that is likely to
		be prohibitively costly and depending on the baseline policy might not be that much more detailed.
	 Whether this kind of combined training actually converges to something is not quite obvious though.)

		The resulting generator could then be used as $\baseline$ with an arbitrary side-effect avoidance
		measure, like relative reachability. For a given state it should be likely to return
		a baseline state corresponding to a possible rollout of the baseline policy from possibly some
		steps in the past.

		\subsubsection{Doubts}
		 Does this actually work?
			\begin{enumerate}
				\item Does training with $L$ actually create a generator that
					generates things mostly corresponding to the $l$th argument with about $2^{-l}$
					probability? Or does this converge to something else, e.g. always the $0$th argument?
				\item Is it reasonable expect that the CVAE model will be able to ``model'' the whole
					transformation needed to get from $s_t$ to $s^{\left( t-4^l \right)}_{4^l}$?
					After all this requires going back $4^l$ steps and back forward using a different policy.
					Even if we had one model per $l$ this might be a nontrivial task. On the other hand this might
					be simpler if the baseline policy and undisturbed environment behaviour is simple.
				\item Variant of the above -- if we use multiple models we could have them generate
					partial (as in only some steps) trajectories. This I at least know is mostly possible
					and might improve accuracy and interpretability.
				\item What about parts of the state space that random policies cannot reach, but the
					actual policy used by the training agent can. How well will this generalize?
			\end{enumerate}
\appendix
	\section{Non-Markovian rewards}
	 In the main text I pretended everything is Markovian, but at least Krakovna's original approach
		is not. A definition that would actually apply to her appraoch is presented below,
		but was too unwieldy to work with and did not actually provide much insight.
		\begin{definition}[Side effect avoidance]
			\emph{Side effect avoidance} is a class of approaches for mitigating reward misspecification, which
			modify the reward passed to the agent according to the equation
			\begin{equation}
				\modifiedRewardFunction\left( \left( s_i \right)_{i=0}^t \right) =
				\corruption\left( s_t \right) -
				\lambda \cdot \deviationFromBaseline\left( \left( s_i \right)_{i=0}^t, \baseline\left( \left( s_i \right)_{i=0}^t \right) \right),
				\label{eq:sideEffectAvoidance-nonMarkov}
			\end{equation}
			where $t$ is the time step, $s_i$ is the state at time $i$, $\baseline$ is a probability
			distribution of trajectories of the baseline agent, and $\deviationFromBaseline$ is a measure
			of the deviation between a trajectory and a baseline distribution of trajectories.
